---
title: "Advanced R Markdown"
author: "BBL"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document:
    df_print: paged
    toc: true
    toc_float: true
    code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(ggplot2)
theme_set(theme_bw())
```

## Topics {data-background=#cceeff}

* Under the hood: YAML, R, Markdown, knitr, pandoc
* R
    + Programmatic reports
* HTML goodies
    + Floating TOC
    + Tabsets
    + Cascading Style Sheets (CSS)
    + Data frames as tables
    + Inline equations
* Other packages
* Interactive figures and maps
  + Plotly
  + Leaflet
* Tables
  + DT
* Citations
* Bookdown

**Goal: exposure to a variety of more advanced R Markdown techniques and tricks.**


## About the YAML header

## HTML goodies

### Printing data frames

For html output you can add the `df_print: paged` parameter to your YAML header to
have printed data frames rendered as html tables.

```
output:
  html_document:
    df_print: paged
```

```{r, mtcars}
mtcars
```


### TOC and code folding

```
---
title: "Advanced R Markdown"
author: "BBL"
date: "`r format(Sys.time(), '%d %B, %Y')`"
date: "`r "\u0060r format(Sys.time(), '%d %B, %Y')\u0060"`"
output:
  html_document:
    toc: true
    toc_float: true
    code_folding: hide
---
```

### Tabs {.tabset}

You can use tabs to organize your content:

```
## Tabs {.tabset}

### Tab 1 name
(content)

### Tab 2 name
(content)
```

#### Cars

```{r}
plot(cars$speed, cars$dist)
```

#### Iris

```{r}
pairs(iris)
```

#### Volcano

```{r}
image(volcano)
```


## Useful R packages

### Interactive graphs

```{r plotly, message=FALSE}
library(plotly)
p <- ggplot(mtcars, aes(hp, mpg, size = cyl, color = disp)) + geom_point()
ggplotly(p)
```


### Interactive tables

```{r DT, message=FALSE}
library(DT)
library(gapminder)
datatable(mtcars, rownames = TRUE, filter = "top", 
          options = list(pageLength = 5, scrollX = TRUE))
```

\
This takes _~~one~~ two_ lines of code in RMarkdown. Example based on [this post](https://holtzy.github.io/Pimp-my-rmd/#use_dt_for_tables).


### Interactive maps

```{r leaflet, out.width='100%'}
library(leaflet)
leaflet() %>% 
  addTiles() %>%
  setView(-76.9219, 38.9709, zoom = 17) %>%
  addPopups(-76.9219, 38.9709,
            "Here is the <b>Joint Global Change Research Institute</b>")
```


## Citations and references

```
In a subsequent paper [@Bond-Lamberty2009-py], we used the same model 
outputs to examine the _hydrological_ implications of these wildfire 
regime shifts [@Nolan2014-us].
```
 In a subsequent paper (Bond-Lamberty et al. 2009), we used the same model outputs to examine the _hydrological_ implications of these wildfire regime shifts (Nolan et al. 2014).

**References**

Bond-Lamberty, Ben, Scott D Peckham, Stith T Gower, and Brent E Ewers. 2009. “Effects of Fire on Regional Evapotranspiration in the Central Canadian Boreal Forest.” Glob. Chang. Biol. 15 (5): 1242–54.

Nolan, Rachael H, Patrick N J Lane, Richard G Benyon, Ross A Bradstock, and Patrick J Mitchell. 2014. “Changes in Evapotranspiration Following Wildfire in Resprouting Eucalypt Forests.” Ecohydrol. 6 (January). Wiley Online Library.


## Resources

Good resources:

* The [R Markdown Cheat Sheet](https://rstudio.com/wp-content/uploads/2015/02/rmarkdown-cheatsheet.pdf)
* The [R Markdown Cookbook](https://bookdown.org/yihui/rmarkdown-cookbook/)
* [15 Tips on Making Better Use of R Markdown](https://slides.yihui.org/2019-dahshu-rmarkdown#1)
* [Stack Overflow](https://stackoverflow.com/questions/tagged/r-markdown) of course
* Allison Hill's [How I Teach R Markdown](https://alison.rbind.io/post/2020-05-28-how-i-teach-r-markdown/)
* [How to Make Beautiful Tables in R](https://rfortherestofus.com/2019/11/how-to-make-beautiful-tables-in-r/)


## The End {data-background=#cceeff}

Thanks for attending this introduction to R Markdown documents workshop! We hope it was useful.

This presentation was made using R Markdown version `r packageVersion("rmarkdown")` running under `r R.version.string`.

These slides are available at https://rpubs.com/bpbond/626346. (And they were written in R Markdown! The code is [here](https://github.com/JGCRI/Rworkshops/blob/master/Introduction_to_RMarkdown.Rmd).)

